Inefficient Concessions and Mediation

### 2016-10-01 draft.tex
update horner, morelli and squintani reference
finish updating reference list for previous citations
revise old intro, remove break for old lit review/model intro secion

This holds for concessions other than land. Russia is loathe to cede further autonomy to Chechnya, even though independence could hypothetically relieve tensions and reduce violence. If Chechen independence could in fact lead to lowered violence, it is possible that this move would be in Russia's best interest. But the Russians perceive, perhaps correctly, that any move towards autonomy might only give Chechens further leverage against the Kremlin. They also realize that were Chechnya to realize its hopes, other ethnic areas like Tatarstan might be emboldened to demand more. From the Russian perspective, granting the concession of increased autonomy would likely have future costs.

In these and many other situations, a lack of trust is an impediment to peace, preventing parties from making necessary concessions. Mediation has long been understood as a mechanism for reducing mistrust, but in most of the published literature on international conflict mediation lacks a rigorous theoretical framework.

This paper has three parts. First, it explores the role of costly concessions in a game theoretic framework, evoking a two-sided version of Spence signaling (1973). Second, after considering existing explanations for inefficient concessions, it proposes a new theoretical explanation for why inefficiency might occur in a conflict setting: inefficient concessions may be preferred if the most efficient concessions are seen as having additional, future costs.\footnote{The type of inefficient concessions addressed here are those whose value to the recipient are lower than the cost to the giver.} The Golan Heights, for example, will not be conceded if Israel foresees future security risks that outweigh the immediate benefits of peace. Finally, this paper uses a mechanism design framework to model a third party's role in bringing about more efficient concessions. Many scholars contend that mediators with enforcement capabilities can help to resolve conflict; this analysis provides an explanation for why enforcement is effective.

\section{Literature review and model introduction}

@@

\item Beardsley, Kyle. (2010). ``Pain, Pressure, and Political Cover: Explaining Mediation Incidence." \emph{Journal of Peace Research}. Forthcoming.

@@

\item Horner, Johannes, Massimo Morelli and Francesco Squintani. (2010). ``Mediation and Peace.'' Cowles foundation. Unpublished manuscript.

>>>>

This holds for concessions other than land. Russia is loathe to cede further autonomy to Chechnya, even though it could hypothetically relieve tensions and reduce violence. If Chechen autonomy could in fact lead to lowered violence, it is possible that this move would be in Russia's best interest. But the Russians perceive, perhaps correctly, that any move towards autonomy might only give Chechens further leverage against the Kremlin. From the Russian perspective, granting the concession of increased autonomy would likely have future costs.

%In these and many other situations, a lack of trust is an impediment to peace, preventing parties from making necessary concessions. Mediation has long been understood as a mechanism for reducing mistrust, but in much of the published literature on international conflict mediation lacks a rigorous theoretical framework.

This paper uses a two-sided Spence signaling (1973) model to propose a new theoretical explanation for why inefficiency might occur in a conflict setting: inefficient concessions may be preferred if the most efficient concessions are seen as having additional, future costs.\footnote{The type of inefficient concessions addressed here are those whose value to the recipient are lower than the cost to the giver.} The Golan Heights, for example, will not be conceded if Israel foresees future security risks that outweigh the immediate benefits of peace. We then suggest a new explanation for the effectiveness of third-party mediation by demonstrating in a mechanism design framework that a mediator can facilitate more efficient concessions.

@@

\item Beardsley, Kyle. (2010). ``Pain, Pressure, and Political Cover: Explaining Mediation Incidence." \emph{Journal of Peace Research}. 47:2, 395-406.

@@

\item Horner, Johannes, Massimo Morelli and Francesco Squintani. (2010). ``Mediation and Peace.'' \emph{The Review of Economic Studies} 84, 1483-1501.
